Community Analysis

### 2022-09-14 R/08-community-analysis.Rmd
update analyses and intermediate data files

@@

all_taxa <-
  all_taxa %>%
	mutate(Family = str_replace(Family, "(.*)", "*\\1*"),
	       Family = str_replace(Family, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))
all_taxa$Family[all_taxa$Family == "Unclassified *0319-6G20*"] <- "Uncultured family 0319-6G20"
all_taxa$Family[all_taxa$Family == "Unclassified *Chloroplast*"] <- "Chloroplast"
all_taxa$Family[all_taxa$Family == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
all_taxa$Family[all_taxa$Family == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
all_taxa$Family[all_taxa$Family == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"

@@

knitr::opts_chunk$set(eval = FALSE)

```{r setup-differential-abundance-data-genus}

aes_Genus_G5 <-
  aes_G5 %>% 
  select(asv, sample, treat, count, Genus) %>%
  group_by(sample, Genus) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  select(sample, treat, Genus, total) %>% 
  distinct()

prevalent_families <- 
  aes_Genus_G5 %>% 
  group_by(Genus, treat) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
  filter(N/12 > 0.1 & P/12 > 0.1) %$%
  Genus

otu_table <- 
  aes_Genus_G5 %>% 
  filter(Genus %in% prevalent_families) %>% 
  select(sample, Genus, total) %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

sample_data <- aes_Genus_G5 %>% 
  select(sample, treat) %>% 
  distinct() %>% 
  as.data.frame()

dt_data <- plot(da_analysis, 
                # level = c("Genus"), 
                data_only = TRUE)

if(file.exists("Output/ancom_out_genus.rds")) {
  out <- 
    readRDS("Output/ancom_out_genus.rds")
} else {
  # Run Ancom
  out = ancom(phyloseq = aes_ps,  
              p_adj_method = "BH",
              prv_cut = 0.1, 
              lib_cut = 0, 
              main_var = "treat",
              rand_formula = NULL, 
              lme_control = NULL,
              struc_zero = TRUE, 
              neg_lb = FALSE, 
              alpha = 0.05, 
              n_cl = 1)  
  out %>% 
    saveRDS("Output/ancom_out_genus.rds")
}

if(file.exists("Output/aldex_x_all_genus.rds")) {
  x.all <- 
    readRDS("Output/aldex_x_all_genus.rds")
} else {
  x.all <- aldex(otu_table, 
                 conds, mc.samples=1000, test="t", effect=TRUE,
                 include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
  x.all %>% 
    saveRDS("Output/aldex_x_all_genus.rds")
}

x.all_effect_1 <-
  x.all %>% 
  filter(abs(effect) > 1) %>% 
  as_tibble(rownames = "Genus")

# x.all_effect_1 %>% 
#   # inner_join(taxa_table, by = "asv")
#   ggplot(aes(y = reorder(Genus, effect), x = effect)) +
#   geom_point()

# x.all_effect_1 %>% 
#   select(Genus, effect)

# ALDEX2
aldex_taxa <- x.all_effect_1$Genus

at_least_one_taxa <-
	aes_Genus_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
	filter(Genus %in% ancom_taxa | Genus %in% aldex_taxa | Genus %in% corncob_taxa) %>%
	mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))

at_least_one_iqm <- 
	at_least_one_taxa %>%
	group_by(treat, Genus) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	at_least_one_taxa %>%
	group_by(Genus) %>%
	summarize(median = median(rel_abund)) %>%
	arrange(median)

at_least_one_taxa$Genus <- factor(at_least_one_taxa$Genus, levels = sort_order$Genus)

at_least_one_taxa %>%
	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
			data = at_least_one_iqm, 
			position = position_dodge(width = 0.75), 
			shape = 21, 
			color = "black") +
	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2'))

```{r}
aldex_effect <-
  x.all_effect_1 %>% 
  select(Genus, effect)

all_taxa <-
	aes_Genus_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
  left_join(aldex_effect, by = "Genus") %>% 
	filter(Genus %in% ancom_taxa & Genus %in% aldex_taxa & Genus %in% corncob_taxa) %>% 
  ungroup()

da_fams <-
  all_taxa %>% 
  select(Genus) %>% 
  distinct() %$%
  Genus

all_taxa <-
  all_taxa %>%
	mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))
# all_taxa$Genus[all_taxa$Genus == "Unclassified *0319-6G20*"] <- "Uncultured Genus 0319-6G20"
# all_taxa$Genus[all_taxa$Genus == "Unclassified *Chloroplast*"] <- "Chloroplast"
# all_taxa$Genus[all_taxa$Genus == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
# all_taxa$Genus[all_taxa$Genus == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
# all_taxa$Genus[all_taxa$Genus == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"

all_iqm <- 
	all_taxa %>%
	group_by(treat, Genus) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	all_taxa %>%
  select(Genus, effect) %>% 
  distinct() %>% 
  arrange(effect)
	# group_by(Genus) %>%
	# summarize(median = median(rel_abund)) %>%
	# arrange(median)

all_taxa$Genus <- factor(all_taxa$Genus, levels = sort_order$Genus)
all_iqm$Genus <- factor(all_iqm$Genus, levels = sort_order$Genus)
all_taxa <-
  all_taxa %>% 
  mutate(direction = if_else(effect > 0, 0, 1))
all_iqm <-
all_taxa %>% 
  select(Genus, direction) %>% 
  distinct() %>% 
  full_join(all_iqm, by = "Genus")
(p <-
	all_taxa %>%
	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
			data = all_iqm, 
			position = position_dodge(width = 0.75), 
			shape = 21, 
			color = "black") +
	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
    facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
theme(
  strip.background = element_blank(),
  strip.text.y = element_blank()
))

ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5)

## All
# ggsave("shared_taxa.png", p)
ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)

# aes_Genus_rclr_plot <-
#   aes_Genus_rclr %>% 
# 	filter(Genus %in% ancom_taxa & Genus %in% aldex_taxa & Genus %in% corncob_taxa) %>% 
#   pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
#   select(Genus, treat, rep, passage, rclr) %>% 
#   distinct() %>% 	
#   mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
# 	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
# 				    "Unclassified *\\1*"))
# 
# aes_Genus_rclr_plot$Genus[aes_Genus_rclr_plot$Genus == "Unclassified *0319-6G20*"] <- "Uncultured Genus 0319-6G20"
# aes_Genus_rclr_plot$Genus[aes_Genus_rclr_plot$Genus == "Unclassified *Chloroplast*"] <- "Chloroplast"
# aes_Genus_rclr_plot$Genus <- factor(aes_Genus_rclr_plot$Genus, levels = sort_order$Genus)
# ggplot(aes_Genus_rclr_plot, aes(x = rclr, y = Genus, color = treat)) +
#   facet_grid( ~ passage) +
#   geom_point(position = position_jitterdodge())

p_effect <-
	all_taxa %>%
	ggplot(aes(x = effect, y = Genus)) +
	geom_point() +

```{r setup-differential-abundance-data-species}

aes_Species_G5 <-
  aes_G5 %>% 
  select(asv, sample, treat, count, Species) %>%
  group_by(sample, Species) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  select(sample, treat, Species, total) %>% 
  distinct()

prevalent_families <- 
  aes_Species_G5 %>% 
  group_by(Species, treat) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
  filter(N/12 > 0.1 & P/12 > 0.1) %$%
  Species

otu_table <- 
  aes_Species_G5 %>% 
  filter(Species %in% prevalent_families) %>% 
  select(sample, Species, total) %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

sample_data <- aes_Species_G5 %>% 
  select(sample, treat) %>% 
  distinct() %>% 
  as.data.frame()

dt_data <- plot(da_analysis, 
                # level = c("Species"), 
                data_only = TRUE)

x.all_effect_1 <-
  x.all %>% 
  filter(abs(effect) > 1) %>% 
  as_tibble(rownames = "Species")

# x.all_effect_1 %>% 
#   # inner_join(taxa_table, by = "asv")
#   ggplot(aes(y = reorder(Species, effect), x = effect)) +
#   geom_point()

# x.all_effect_1 %>% 
#   select(Species, effect)

# ALDEX2
aldex_taxa <- x.all_effect_1$Species

at_least_one_taxa <-
	aes_Species_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
	filter(Species %in% ancom_taxa | Species %in% aldex_taxa | Species %in% corncob_taxa) %>%
	mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))

at_least_one_iqm <- 
	at_least_one_taxa %>%
	group_by(treat, Species) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	at_least_one_taxa %>%
	group_by(Species) %>%
	summarize(median = median(rel_abund)) %>%
	arrange(median)

at_least_one_taxa$Species <- factor(at_least_one_taxa$Species, levels = sort_order$Species)

at_least_one_taxa %>%
	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
			data = at_least_one_iqm, 
			position = position_dodge(width = 0.75), 
			shape = 21, 
			color = "black") +
	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2'))

```{r}
aldex_effect <-
  x.all_effect_1 %>% 
  select(Species, effect)

all_taxa <-
	aes_Species_G5 %>%
	group_by(sample) %>%
	mutate(rel_abund = (total / sum(total)) * 100) %>%
  left_join(aldex_effect, by = "Species") %>% 
	filter(Species %in% ancom_taxa & Species %in% aldex_taxa & Species %in% corncob_taxa) %>% 
  ungroup()

da_fams <-
  all_taxa %>% 
  select(Species) %>% 
  distinct() %$%
  Species

all_taxa <-
  all_taxa %>%
	mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))
all_taxa$Species[all_taxa$Species == "Unclassified *0319-6G20*"] <- "Uncultured Species 0319-6G20"
all_taxa$Species[all_taxa$Species == "Unclassified *Chloroplast*"] <- "Chloroplast"
all_taxa$Species[all_taxa$Species == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
all_taxa$Species[all_taxa$Species == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
all_taxa$Species[all_taxa$Species == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"

all_iqm <- 
	all_taxa %>%
	group_by(treat, Species) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")

sort_order <-
	all_taxa %>%
  select(Species, effect) %>% 
  distinct() %>% 
  arrange(effect)
	# group_by(Species) %>%
	# summarize(median = median(rel_abund)) %>%
	# arrange(median)

all_taxa$Species <- factor(all_taxa$Species, levels = sort_order$Species)
all_iqm$Species <- factor(all_iqm$Species, levels = sort_order$Species)
all_taxa <-
  all_taxa %>% 
  mutate(direction = if_else(effect > 0, 0, 1))
all_iqm <-
all_taxa %>% 
  select(Species, direction) %>% 
  distinct() %>% 
  full_join(all_iqm, by = "Species")
(p <-
	all_taxa %>%
	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
			data = all_iqm, 
			position = position_dodge(width = 0.75), 
			shape = 21, 
			color = "black") +
	theme_bw() +
	theme(axis.text.y = element_markdown()) +
	labs(x = "Relative Abundance (%)", y = NULL) +
	scale_color_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
	scale_fill_manual(name = "Treatment",
			   labels = c('Neutral', 'Positive'),
			   values = c('gray40', 'darkorange2')) +
    facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
theme(
  strip.background = element_blank(),
  strip.text.y = element_blank()
))

ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5)

## All
# ggsave("shared_taxa.png", p)
ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)

# aes_Species_rclr_plot <-
#   aes_Species_rclr %>% 
# 	filter(Species %in% ancom_taxa & Species %in% aldex_taxa & Species %in% corncob_taxa) %>% 
#   pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
#   select(Species, treat, rep, passage, rclr) %>% 
#   distinct() %>% 	
#   mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
# 	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
# 				    "Unclassified *\\1*"))
# 
# aes_Species_rclr_plot$Species[aes_Species_rclr_plot$Species == "Unclassified *0319-6G20*"] <- "Uncultured Species 0319-6G20"
# aes_Species_rclr_plot$Species[aes_Species_rclr_plot$Species == "Unclassified *Chloroplast*"] <- "Chloroplast"
# aes_Species_rclr_plot$Species <- factor(aes_Species_rclr_plot$Species, levels = sort_order$Species)
# ggplot(aes_Species_rclr_plot, aes(x = rclr, y = Species, color = treat)) +
#   facet_grid( ~ passage) +
#   geom_point(position = position_jitterdodge())

p_effect <-
	all_taxa %>%
	ggplot(aes(x = effect, y = Species)) +
	geom_point() +

troph_genus_data <-
	aes_data %>%
	filter(Family %in% methanotrophs) %>%
	select(sample, count, Genus, passage, treat) %>%
	group_by(sample, Genus) %>%
	mutate(total = sum(count)) %>%
	select(-count) %>%
	ungroup() %>%
	group_by(sample) %>%
	mutate(rel_abund = total / sum(total)) %>%
	ungroup() %>%
	distinct() 
troph_genus_iqm <- 
	troph_genus_data %>%
	group_by(passage, treat, Genus) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")
troph_genus_data %>%
	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	facet_grid( ~ passage) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_genus_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
	theme_bw()
ggsave("troph_genus.png")

aes_data %>%
	select(sample, count, Genus, passage, treat) %>%
	group_by(sample, Genus) %>%
	mutate(total = sum(count)) %>%
	select(-count) %>%
	ungroup() %>%
	group_by(sample) %>%
	mutate(rel_abund = total / sum(total)) %>%
	ungroup() %>%
	filter(Genus %in% c("Methylocella", "Microvirga", "Methylorosula")) %>%
	distinct() %>%
	ggplot(aes(x = rel_abund, y = Genus)) +
	geom_jitter() +
	facet_grid(treat ~ passage)

troph_family_data <-
	aes_data %>%
	select(sample, count, Family, passage, treat) %>%
	group_by(sample, Family) %>%
	mutate(total = sum(count)) %>%
	select(-count) %>%
	ungroup() %>%
	group_by(sample) %>%
	mutate(rel_abund = total / sum(total)) %>%
	ungroup() %>%
	filter(Family %in% methanotrophs) %>%
	distinct()
troph_family_iqm <- 
	troph_family_data %>%
	group_by(passage, treat, Family) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")
troph_family_data %>%
	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	facet_grid( ~ passage) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_family_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
	theme_bw()
ggsave("troph_family.png")

troph_species_data <-
	aes_data %>%
	filter(Family %in% methanotrophs) %>%
	select(sample, count, Species, passage, treat) %>%
	group_by(sample, Species) %>%
	mutate(total = sum(count)) %>%
	select(-count) %>%
	ungroup() %>%
	group_by(sample) %>%
	mutate(rel_abund = total / sum(total)) %>%
	ungroup() %>%
	distinct()
troph_species_iqm <- 
	troph_species_data %>%
	group_by(passage, treat, Species) %>%
	summarize(median = median(rel_abund), 
		  first_quartile = quart(rel_abund)[1],
		  third_quartile = quart(rel_abund)[2], .groups = "drop")
troph_species_data %>%
	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
	geom_point(position=position_jitterdodge()) +
	facet_grid( ~ passage) +
	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_species_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
	theme_bw()
ggsave("troph_species.png")

```{r aldex-methanotrophs-family}

x.all <- aldex(troph_data_family, 
               conds, mc.samples=1000, test="t", effect=TRUE,
     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)

x.all

```{r aldex-methanotrophs-genus}

x.all <- aldex(troph_data_genus, 
               conds, mc.samples=1000, test="t", effect=TRUE,
     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)

x.all

```{r}
## ANCOM
```
```{r}

# Setup phyloseq objects

ps_family <- phyloseq(otu_table(troph_data_family,
                                taxa_are_rows = TRUE),
                      sample_data(sample_data))
ps_genus <- phyloseq(otu_table(troph_data_genus,
                                taxa_are_rows = TRUE),
                      sample_data(sample_data))

# Run Ancom

out_family = ancom(phyloseq = ps_family,  
                  p_adj_method = "BH",
                  prv_cut = 0.1, 
                  lib_cut = 0, 
                  main_var = "treat",
                  rand_formula = NULL, 
                  lme_control = NULL,
                  struc_zero = TRUE, 
                  neg_lb = FALSE, 
                  alpha = 0.05, 
                  n_cl = 1)

out = ancom(phyloseq = ps_genus,  
                  p_adj_method = "BH",
                  prv_cut = 0.1, 
                  lib_cut = 0, 
                  main_var = "treat",
                  rand_formula = NULL, 
                  lme_control = NULL,
                  struc_zero = TRUE, 
                  neg_lb = FALSE, 
                  alpha = 0.05, 
                  n_cl = 1)

q_val = out$q_data
beta_val = out$beta_data
# Only consider the effect sizes with the corresponding q-value less than alpha
beta_val = beta_val * (q_val < 0.05) 
# Choose the maximum of beta's as the effect size
beta_pos = apply(abs(beta_val), 2, which.max) 
beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
                  FUN.VALUE = double(1))

# Number of taxa except structural zeros
n_taxa = ifelse(is.null(out$zero_ind), 
                nrow(feature_table), 
                sum(apply(out$zero_ind, 1, sum) == 0))

# Cutoff values for declaring differentially abundant taxa
cut_off = 0.7 * (n_taxa - 1)

df_fig_w = out$res %>%
  dplyr::mutate(beta = beta_max,
                direct = case_when(
                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
                  TRUE ~ "Not Significant"
                  )) %>%
  dplyr::arrange(W)
df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
df_fig_w$direct = factor(df_fig_w$direct, 
                     levels = c("Negative", "Positive", "Not Significant"))

otu_table <-  aes_data %>% 
  filter(asv %in% troph_asvs) %>% 
  filter(passage == "G5") %>% 
  select(sample, count, Genus) %>% 
  group_by(Genus, sample) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  mutate(treat = substr(sample, 3, 3)) %>% 
  group_by(Genus, treat) %>% 
  mutate(prev = sum(total > 0)) %>% 
  ungroup() %>% 
  filter(prev > 2) %>% 
  select(-treat, -prev) %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

rownames(otu_table) <- otu_table[, 1]
otu_table <- otu_table[, -1]

conds <- 
  otu_table %>% 
  colnames() %>% 
  substr(3, 3)

x.all <- aldex(otu_table, 
               conds, mc.samples=1000, test="t", effect=TRUE,
     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)

x.all_effect_1 <-
  x.all %>% 
  filter(abs(effect) > 0.5) %>% 
  as_tibble(rownames = "asv")

x.all_effect_1 %>% 
  ggplot(aes(y = reorder(asv, effect), x = effect)) +
  geom_point()

```{r aldex_genus_everything}

otu_table <- 
  aes_data %>% 
  filter(passage == "G5") %>% 
  select(Genus, count, sample) %>% 
  group_by(Genus, sample) %>% 
  summarize(total = sum(count), .groups = "drop") %>% 
  mutate(treat = substr(sample, 3, 3)) %>% 
  group_by(Genus, treat) %>% 
  mutate(prev = sum(total > 0)) %>% 
  ungroup() %>% 
  filter(prev > 2) %>% 
  select(-treat, -prev) %>% 
  pivot_wider(names_from = "sample", values_from = "total", values_fill = 0) %>% 
  as.data.frame()

rownames(otu_table) <- otu_table[, 1]

otu_table <- otu_table[, -1]

x.all <- aldex(otu_table, 
               conds, mc.samples=1000, test="t", effect=TRUE,
     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)

x.all_effect_1 <-
  x.all %>% 
  filter(abs(effect) > 1) %>% 
  as_tibble(rownames = "Genus")

x.all_effect_1 %>% 
  ggplot(aes(y = reorder(Genus, effect), x = effect)) +
  geom_point()

```{r ancova-clr}

gm <- function(x){

  exp(mean(log(x[x>0])))

rclr <- function(x) {
  log(x/gm(x))
}

aes_family <-
  aes_data %>%
  select(Family, count, sample, passage, treat) %>%
  group_by(Family, sample) %>%
  mutate(total = sum(count)) %>%
  select(-count) %>%
  ungroup() %>%
  distinct() %>%
  mutate(rep = as.numeric(substr(sample, 4, 5))) %>%
  select(-sample)

prevalent_families <- 
  aes_family %>% 
  mutate(passagetreat = paste0(passage, treat)) %>%
  group_by(Family, passagetreat) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "passagetreat", values_from = "prev", values_fill = 0) %>% 
  filter(G2N/12 > 0.1 & G2P/12 > 0.1 & G5N/12 > 0.1 & G5P/12 > 0.1) %$%
  Family

aes_family_rclr <-
  aes_family %>% 
  filter(Family %in% prevalent_families) %>% 
  group_by(passage, treat, rep) %>%
  mutate(rclr = log(total/gm(total))) %>%
  select(-total) %>%
  ungroup() %>%
  pivot_wider(names_from = passage, values_from = rclr)

ancova <- function(df) {
  lm(G5 ~ G2 + treat, data = df, na.action = )
}

model_results <-
aes_family_rclr %>%
  nest(data = -Family) %>%
  mutate(model = map(data, ancova)) %>%
  mutate(tidy = map(model, tidy)) %>% 
  unnest(tidy) %>% 
  filter(term == "treatP")

model_results$p.adjust <- p.adjust(model_results$p.value, method = "BH")

model_results %>%
  filter(p.adjust < 0.05) %>% 
  arrange(estimate) %>% 
  print(n = 22)

model_results %>% 
  filter(p.adjust < 0.05) %>% 
  arrange(estimate) %>% 
  unnest(data) %>% 
  ggplot(aes(x = estimate, xmax = estimate + std.error, xmin = estimate - std.error, y = reorder(Family, estimate))) +
  geom_pointrange() +
  labs(x = "Estimate \n(log-ratio increase in P)", y = NULL) +
  theme_bw()
ggsave("ancova-plot.png", width = 5, height = 5)

model_results %>% 
  filter(p.adjust < 0.05) %>% 
  unnest(data) %>% 
  pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
  select(Family, treat, rep, passage, rclr) %>% 
  distinct() %>% 
  ggplot(aes(x = rclr, y = Family, color = treat)) +
  facet_grid( ~ passage) +
  geom_point(position = position_jitterdodge())

aes_family_data <-
  aes_data %>% 
  select(asv, sample, passage, treat, count, Family) %>%
  group_by(sample, Family) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  select(sample, treat, passage, Family, total) %>% 
  distinct()

prevalent_families <- 
  aes_family_data %>% 
  group_by(Family, treat, passage) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
  filter(N/12 > 0.1 & P/12 > 0.1) %$%
  Family

aes_family_data <-
  aes_family_data %>% 
  filter(Family %in% prevalent_families) %>% 
  mutate(jar = substr(sample, 3, 5)) %>%
  select(-sample) %>%
  pivot_wider(names_from = "passage", values_from = "total", values_fill = 0)

ancova_model <- function(df) {
	lm(G5 ~ G2 + treat, data = df)
}

aes_family_nested <-
	aes_family_data %>%
	group_by(Family) %>%
	nest() %>%
	mutate(model = map(data, ancova_model)) %>%
	mutate(tidy = map(model, tidy)) %>%
	ungroup()

aes_ancova <- aes_family_nested %>%
	unnest(tidy) %>%
	filter(term == "treatP")

aes_ancova$p.adjust <- p.adjust(aes_ancova$p.value, method = "BH")

aes_ancova %>%
	arrange(estimate) %>%
	filter(p.adjust < 0.05) %>%
	print(n = 22)

aes_family_data <-
  aes_data %>% 
  select(asv, sample, passage, treat, count, Family) %>%
  group_by(sample, Family) %>% 
  mutate(total = sum(count)) %>% 
  ungroup() %>% 
  select(sample, treat, passage, Family, total) %>% 
  distinct()

prevalent_families <- 
  aes_family_data %>% 
  group_by(Family, treat, passage) %>% 
  summarize(prev = sum(total > 0), .groups = "drop") %>% 
  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
  filter(N/12 > 0.1 & P/12 > 0.1) %$%
  Family

otu_table <- 
  aes_family_data %>% 
  filter(Family %in% prevalent_families) %>% 
  select(sample, Family, total) %>% 
  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
  as.data.frame()

sample_data <- aes_family_data %>% 
  select(sample, treat, passage) %>% 
  distinct() %>% 
  as.data.frame()

rownames(sample_data) <- sample_data[,1]
sample_data <- sample_data[, -1]
sample_data <-
	sample_data %>%
	mutate(passage12 = if_else(passage == "G2", 1, 2))

set.seed(123)
out2 = ancom(phyloseq = aes_ps,  p_adj_method = "holm",
             prv_cut = 0.10, lib_cut = 1000, main_var = "treat", 
             adj_formula = "passage12", 
             rand_formula = "~ passage12 | 1",
             lme_control = list(maxIter = 100, msMaxIter = 100, opt = "optim"), 
             struc_zero = TRUE, neg_lb = FALSE, alpha = 0.05, n_cl = 2)

res2 = out2$res

out2 = ancom(phyloseq = family_data2,  p_adj_method = "holm",
             prv_cut = 0.10, lib_cut = 1000, main_var = "group", 
             adj_formula = "nationality + timepoint + bmi_group", 
             rand_formula = "~ timepoint | subject",
             lme_control = list(maxIter = 100, msMaxIter = 100, opt = "optim"), 
             struc_zero = TRUE, neg_lb = TRUE, alpha = 0.05, n_cl = 2)

res2 = out2$res

x.all_effect_1$Family

dt_data$taxa

```{r}
combined_da <-
  dt_data %>% 
  as_tibble %>% 
  full_join(sig_families, by = c("taxa" = "Family")) %>% 
  full_join(x.all_effect_1, by = c("taxa" = "Family")) %>% 
  full_join(df_fig_w[df_fig_w$detected_0.7,], by = c("taxa" = "taxon_id"))

shared_da <-
  dt_data %>% 
  as_tibble %>% 
  inner_join(sig_families, by = c("taxa" = "Family")) %>% 
  inner_join(x.all_effect_1, by = c("taxa" = "Family")) %>% 
  inner_join(df_fig_w[df_fig_w$detected_0.7,], by = c("taxa" = "taxon_id"))

shared_da %>% 
  ggplot(aes(y = reorder(taxa, effect), x = effect)) +
  geom_point()

>>>>

```{r remove_chloroplasts, eval=FALSE}

chloroplast_seqs <- 
  aes_data %>% 
  filter(Order == 'Chloroplast') %>% 
  select(asv, seq) %>% 
  distinct() %$%
  seq
# 10 ASVs were identified as chloroplast in their taxonomy.
# seq matches complete genomes of chloroplasts on NCBI so I will remove them.

aes_data <- 
  aes_data %>% 
  filter(Order != 'Chloroplast')

@@

all_taxa <-
  all_taxa %>%
	mutate(Family = str_replace(Family, "(.*)", "*\\1*"),
	       Family = str_replace(Family, "\\*(.*)_unclassified\\*",
				    "Unclassified *\\1*"))
all_taxa$Family[all_taxa$Family == "Unclassified *0319-6G20*"] <- "Uncultured family 0319-6G20"
# all_taxa$Family[all_taxa$Family == "Unclassified *Chloroplast*"] <- "Chloroplast"
all_taxa$Family[all_taxa$Family == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
all_taxa$Family[all_taxa$Family == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
all_taxa$Family[all_taxa$Family == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"

@@
